---
title: "Getting started with roller"
author: "Nabeil Kizilbash"
date: "December 2, 2018"
output: html_document
vignette: >
  %\VignetteIndexEntry{Introduction to roller}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, echo = FALSE, message = FALSE}
knitr::opts_chunk$set(collapse = T, comment = "#>")
library(roller)
```


## Rolling a Die

The package `"roller"` is a minimal implementation for simulating rolling
a die a given number of times, and to visualize the relative frequencies
in such a series of rolls

## Creating a `device`

The first step is to create a `"die"` object with the function `device()`:

```{r}
mydie <- device()
mydie
```

By default, `device()` creates a die with two sides `1`, and 
`2`, and with probability values for each side of 0.5.

You can change the defatuls with arguments `sides` and `prob`:

```{r}
loaded <- device(sides = c('HEADS', 'tails'), prob = c(0.8, 0.2))
loaded
```


## rolling a die

Once you have defined a `"die"` object, you can use the function `roll()` to
roll a die a given number of `times`. The output will be an object of class 
`"rolls"`, which will contain the vector of `rolls`:

```{r}
set.seed(123)
roll100 <- roll(mydie, times = 100)
roll100
```

The `summary()` function gives you a nice table of frequencies:
```{r}
summary(roll100)
```


## Plot of relative frequencies

You can use the function `plot()` to create a graohic with the relative 
frequencies of a series of coin tosses:

```{r, fig.show='hold'}
plot(roll100)
```

## Additional things you can do

you can extract a specific roll from your rolls using the "[]" operator
```{r}
roll100[17]
```

You can replace a particular roll value with another value
```{r}
roll100[17] <- 2
roll100[17]
```
You can add more rolls via the "+" method
```{r}
roll200 <- (roll100 + 100)

length(roll100$rolls)
length(roll200)

```


