---
title: "Lab 9: Testing functions, and string basics"
subtitle: "Stat 133, Fall 2018"
author: "Nabeil Kizilbash"
output: html_document
fontsize: 11pt
urlcolor: blue
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, error = TRUE)
library(testthat)
```




#Part 2, Basics of String Manipulation
```{r}

head(USArrests)

states <- rownames(USArrests)
head(states)
nchar(states)

```


```{r}
tolower(states)
toupper(states)
casefold(states)


num_chars <- nchar(states)
num_chars

char_freqs<- table(num_chars)




barplot(char_freqs, xlab = "Number of letters",ylab = "frequency" )

```

## Pasting Strings

```{r}
paste('pumpkin', 'pie')
```
You can give it any number of vector inputs

```{r abcde}
paste('a', 'b', 'c', 'd', 'e')
```

You can change the separator with `sep`

```{r abcde-sep}
paste('a', 'b', 'c', 'd', 'e', sep = '-')
```

`paste()` is vectorized:

```{r vectorized}
paste('a', 1:5, sep = '.')
```
```{r paste0}
# paste0() -vs- paste(..., sep = "")
paste0('Pumpkin', 'Pie')
paste('Pumpkin', 'Pie', sep = '')
# paste0() is also vectorized
paste0('a', 1:5)
```




```{r}
head(num_chars)

paste(head(states,5), head(num_chars,5), sep =  " = ")

paste(head(states,5), collapse = "")
```


#Permutations

210 Permutations can be formed using all the given letters in the word Alabama. 7!/4! = 210


#Substrings

```{r}
# extract first 3 characters
substr('Berkeley', 1, 3)

substr(states, 1, 3)

substr(states, nchar(states) - 2, nchar(states))


```
#How would you shorten the state names using the first letter and the last 3-letters? 
What I would do is i would first extract the first letter, then the last 3 letters, then concatonate them together so that the full string is formed.

### Challenge

We already obtained a frequency table `char_freqs` with the counts of state 
names by number of characters. You can use those frequencies to get those 
state-names with 4-characters or 10-characters:

```{r}
# 4-char states
states[num_chars == 4]
```

```{r}
# 10-char states
states[num_chars == 10]
```

You can use `paste()` to join the 4-character states in one single string 
(i.e. _collapsing_) like this---separated by a comma and space---:

```{r}
# collapse 4-char states
paste(states[num_chars == 4], collapse = ", ")
```


## Converting from Fahrenheit Degrees

Here are four functions that convert from Fahrenheit degrees to other temperature scales:

```{r}
to_celsius <- function(x = 1) {
  (x - 32) * (5/9)
}
to_kelvin <- function(x = 1) {
  (x + 459.67) * (5/9)
}
to_reaumur <- function(x = 1) {
  (x - 32) * (4/9)
}
to_rankine <- function(x = 1) {
  x + 459.67
}
```

We can use the previous functions to create a more general function `temp_convert()`:

```{r}
temp_convert <- function(x = 1, to = "celsius") {
  switch(tolower(to),
         "celsius" = to_celsius(x),
         "kelvin" = to_kelvin(x),
         "reaumur" = to_reaumur(x),
         "rankine" = to_rankine(x))
}
temp_convert(30, 'celsius')
```

`temp_convert()` works fine when the argument `to = 'celsius'`. But what happens 
if you try `temp_convert(30, 'Celsius')` or `temp_convert(30, 'CELSIUS')`?

__Your turn__. Rewrite `temp_convert()` such that the argument `to` can be given 
in upper or lower case letters. For instance, the following three calls should 
be equivalent:

MOdify the function to take any input and convert it to lowercase.

```{r, eval=FALSE}
temp_convert(30, 'celsius')
temp_convert(30, 'Celsius')
temp_convert(30, 'CELSIUS')
```


## Names of files

Imagine that you need to generate the names of 10 data `.csv` files.
All the files have the same prefix name but each of them has a different number: 
`file1.csv`, `file2.csv`, ... , `file10.csv`. 

How can you generate a character vector with these names in R? Come up with at 
least three different ways to get such a vector:
```{r}
#Use the substring function to select everything except the last 3 characters, so everything except for the .csv portion of the filenmae

#Extract just the number by excluding the file portion of the file names.

#Do a for loop and only take the part of the filenamethat is a number, that way you are only getting the unique part of each filename


```

Now imagine that you need to rename the characters `file` into `dataset`. In 
other words, you want the vector of file names to look like this: 
`dataset1.csv`, `dataset2.csv`, ... , `dataset10.csv`. Take the previous vector 
of file names and rename its elements:

```{r}
# rename vector of file names by using the substr function to replace the characters at the beginning of the string with dataset.
```



## Using function `cat()`

Run the following code:

```{r, eval=FALSE}
# name of output file
outfile <- "output.txt"
# writing to 'outfile.txt'
cat("title: Lab 09
author: Nabeil Kizilbash
date: 10/27/2018
output: html_document", file = outfile)
# insert new line
cat("\n", file = outfile, append = TRUE)
cat("A 2nd line", file = "output.txt", append = TRUE)
# insert 2 new lines
cat("\n\n", file = outfile, append = TRUE)
cat("\nThe quick brown fox jumps over the lazy dog\n",
    file = outfile, append = TRUE)
```





## Valid Color Names

The function `colors()` returns a vector with the names (in English) of 657 
colors available in R. Write a function `is_color()` to test if a given 
name---in English---is a valid R color. If the provided name is a valid R color, 
`is_color()` returns `TRUE`.  If the provided name is not a valid R color 
`is_color()` returns `FALSE`.

```{r, eval=FALSE}
# your is_color() function
vec <- colors()


#' Color checker
#'
#' @param color 
#'
#' @return true or false
#' @purpose checks whether or not a function is a color in r
is_color <- function(color){
  
  if(color %in% vec){ TRUE}
  
  else{FALSE}
  
  
  
  
}
# test it:
is_color('yellow')  # TRUE
is_color('blu')     # FALSE
is_color('turkuoise') # FALSE
```

## Plot with a valid color

Use `is_color()` to create the function `colplot()` that takes one argument 
`col` (the name of a color) to produce a simple scatter plot with random
numbers (e.g. use `runif()` or `rnorm()` to get point coordinates).

If `col` is a valid name---say "blue"---, the scatterplot should show a title 
`"Testing color blue"`. 

If the provided `col` is not a valid color name, e.g. "blu", then the function 
must stop, showing an __error message__ `"invalid color blu"`.

```{r eval = FALSE}

colplot <- function(x){
  
  if(is_color(x)){
    plot(rnorm(100), col = x, main = "testing color")
  }
  
  else {stop("This is not a valid color")}
  
}



# this should plot
colplot('tomato')




# this stops with error message
colplot('tomate')
```

-----










